Inefficient Concessions and Mediation

### 2019-10-26 presentation.tex
change date on presentation for CesIFO session
compile handout version of presentation (fix typo) for integration into paper
update presentation for CPEG

\begin{document}
\title[Inefficient Concessions and Mediation\hspace{2.95in}\insertframenumber/\inserttotalframenumber]{Inefficient Concessions and Mediation}
\author[Kristy Buzard and Ben Horne]{Kristy Buzard and Ben Horne\\ kbuzard@syr.edu}
\date{June 15, 2017}
\maketitle
%\insertpresentationendpage removed b/c of appendix

@@

\pause
Start with simple, two-player repeated Prisoners' Dilemma
\pause
\begin{itemize}[<+->]
	\item Asymmetric information about $\de_i$: can partner commit to peace?
\end{itemize}

@@

\begin{frame}{Literature}
\pause
Signaling: Spence (1973), but signal has material value
\pause
		\begin{itemize}
			\item cost/benefit tradeoff differs b/c of $\de_i$, $C_i(g)$
		\end{itemize}

@@

\section{Benchmark}
\subsection{No Money Burning}

@@

\pause
\vskip.2in
Some equilibria of interest
\pause
\begin{itemize}[<+->]
	\item Separating without concessions
	\item Separating through concessions
	\item Pool on `Fight'
\end{itemize}
\end{frame}

@@

\pause	
\[
  U_h=pg - c_h(g) + \textstyle \frac{\de_h}{(1-\de_h)} \left[pT+(1-p)(W-D)\right]
\]
\pause
\vskip-.1in

	\begin{itemize}[<+->]
		\item The smallest concession is $p(T+D)$
		\item If $p$ is low, high types are better off in the `fight' pooling equilibrium
	\end{itemize}
\end{frame}

\section{Inefficient Concessions}
\subsection{Add Money Burning}
\begin{frame}{Timeline}
\pause
\begin{enumerate}[<+->]
	\item[-1.] Nature independently determines types of Country 1 and Country 2: $\de_i \in \left\{\de_h,\de_l\right\}$
	\item[0.] Concessions
		\begin{enumerate}
			\item[0a.] Countries simultaneously give costly concessions and decide on the efficiency of the concessions: $(g_i, e_i) \in \mathbb{R}_+ \times [0,1]$ 
			\item[0b.] \textcolor{red!75!green!50!blue!25!gray}{Countries decide what proportion of a received concession to invest in civil society (vs. military capabilities): $\alpha_i \in [0,1]$}
		\end{enumerate}	
	\item[1--$\infty$.] Countries engage in a simultaneous Prisoners' Dilemma interaction 
\end{enumerate}
\end{frame}

\begin{frame}{Burning money is unattractive}
\pause
Now benefit of concession is $eg$

\vskip.1in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 3}
  When concessions confer no material help or harm on the giver, it is optimal to give efficient gifts ($e=1$) in a separating eqm.
\end{beamerboxesrounded}
\pause
\begin{itemize}[<+->]
	\item The benefit of the gift appears on both sides of the incentive constraint for both individuals, so cancels out
	\item Costs of giving a concession don't change
	\item The benefit appears in the high type's expected utility
\end{itemize}
\end{frame}

\begin{frame}{Timeline}
\pause
\begin{enumerate}[<+->]
	\item[-1.] Nature independently determines types of Country 1 and Country 2: $\de_i \in \left\{\de_h,\de_l\right\}$
	\item[0.] Concessions
		\begin{enumerate}
			\item[0a.] Countries simultaneously give costly concessions and decide on the efficiency of the concessions: $(g_i, e_i) \in \mathbb{R}_+ \times [0,1]$ 
			\item[0b.] \textcolor{blue}{Countries decide what proportion of a received concession to invest in civil society (vs. military capabilities): $\alpha_i \in [0,1]$}
		\end{enumerate}	
	\item[1--$\infty$.] Countries engage in a simultaneous Prisoners' Dilemma interaction 
\end{enumerate}
\end{frame}

\begin{frame}[label=hurt]
\frametitle{Concessions can hurt the giver}
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 4}
  When the low type country can use concessions to reduce the payoffs of its negotiating partner during a `fight' stage (i.e. $\alpha_L < 1$), there are parameters under which the separating-through-concessions equilibrium features inefficient concessions.
\end{beamerboxesrounded}
\hyperlink{Supplemental}{\beamerbutton{Modified Payoffs}}
\pause
\[
  U_h=peg - c_h(g) + \textstyle \frac{\de_h}{1-\de_h} \left[pT(1+eg)+(1-p)(W-D(1+eg))\right]
\]
\pause
\vskip-.1in
\begin{itemize}[<+->]
	\item If $p$ is low, concessions likely to be used against you, so remove material value
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Peace not possible in some scenarios}
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 5}
  When $\alpha_L < 1$ is allowed, peace becomes unachievable for some parameters under which it is achievable when concessions do not have material value.
\end{beamerboxesrounded}

\pause
\vskip.1in
\begin{itemize}[<+->]
	\item If peace is achievable, concessions may be either efficient or inefficient
	\item High-type utility may increase or decrease from no-material-value case
\end{itemize}
\end{frame}

\section{Mediation}
\subsection{}
\begin{frame}{Mediator as mechanism designer}
\pause
`Manipulative' mediator: parties report their types, must deliver the stipulated concessions
\pause
\begin{itemize}[<+->]
  \item Mechanism: if two high types, give concession and play `Trust'; Otherwise, no concession and `Fight'
	\item Concession is necessary to get truthful revelation, but only to high type
	\item Need cost of concession for low type to be not to large relative to cost for high type
\end{itemize}
\vskip.1in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 6}
  A mediator restores peace where $\alpha_i < 1$ destroys it, eliminates inefficient concessions, and reduces the threshold $\de_h$ for high types.
\end{beamerboxesrounded}

@@

Back to \hyperlink{hurt}{\beamerbutton{Concessions can hurt the giver}}.
\end{frame}

>>>>

\begin{document}
\title[Inefficient Concessions and Mediation\hspace{2.95in}\insertframenumber/\inserttotalframenumber]{Inefficient Concessions and Mediation}
\author[Kristy Buzard and Ben Horne]{Kristy Buzard and Ben Horne\\ kbuzard@syr.edu}
\date{October 25, 2019}
\maketitle
%\insertpresentationendpage removed b/c of appendix

@@

\pause
Start with simple, two-player repeated Prisoners' Dilemma
\pause
\begin{itemize}[<+->]
	\item Asymmetric information about partner's discount factor $\de_i$: can partner commit to peace?
\end{itemize}

@@

\begin{frame}{Literature}
\pause
Signaling: Spence (1973), but signal has potential future cost
\pause
		\begin{itemize}
			\item cost/benefit tradeoff b/c of $\de_i$, $C_i(g)$
		\end{itemize}

@@

\section{Benchmark}
\subsection{No Money Burning, No Impact on Giver}

@@

\pause
\vskip.2in
Some equilibria of interest
\pause
\begin{itemize}[<+->]
	\item Pool on `Fight'
	\item Separating through concessions
	\item Separating without concessions
\end{itemize}
\end{frame}

\begin{frame}{Pooling Equilibrium}
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Lemma 4}
  From period 1 on, playing fight in all periods is the only sequentially rational strategy for low types regardless of their beliefs of the other country's type and strategy.
\end{beamerboxesrounded}

\pause
\vskip.1in

	\begin{itemize}[<+->]
		\item This equilibrium can always be chosen by both types
	\end{itemize}
\end{frame}

@@

\pause	
\vskip.1in
If $p$ is low enough, high types are better off in the `fight' pooling equilibrium
\pause
\[
  U_h(CSE)=pg - g + \textstyle \frac{1}{(1-\de_h)} \left[pT+(1-p)(W-D)\right]
\]
\pause
\[
  U_h(POOL)=\textstyle \frac{1}{(1-\de_h)} \left(W-D\right)
\]

\begin{frame}{Smallest Separating Concession}
The smallest concession, $p(T+D)$, comes from the low type's incentive constraint
\pause
\[
  pg +X_{FF}^l \geq -g + pg + pX_{FT}^l+(1-p)X_{FF}^l 
\]
\pause
\[
  pg +\frac{W-D}{1-\de_l} \geq -g + pg + p\left[T+W + \frac{\de \left(W-D\right)}{1-\de_l} \right]+(1-p)\frac{W-D}{1-\de_l}
\]
\pause
\[
  g \geq p\left[T+W + \frac{\de \left(W-D\right)}{1-\de_l} \right]-p\frac{\left(W-D\right)}{1-\de_l}
\]
\pause
\[
  g \geq pT+pW -pW -pD + p\frac{\left(W-D\right)}{1-\de_l} -p\frac{\left(W-D\right)}{1-\de_l}
\]
\end{frame}

\section{Inefficient Concessions}
\subsection{Add Impact of Concessions on Giver (Still no money burning)}
\begin{frame}[label = hurt]{Timeline}
\pause
\begin{enumerate}[<+->]
	\item[-1.] Nature independently determines types of Country 1 and Country 2: $\de_i \in \left\{\de_h,\de_l\right\}$
	\item[0.] Concessions
		\begin{enumerate}
			\item[0a.] Countries decide what proportion of a received concession to invest in civil society (vs. military capabilities): $\alpha_i \in [0,1]$ \hskip.2in \hyperlink{Supplemental}{\beamerbutton{Modified Payoffs}}
			\item[0b.] \textcolor{red!75!green!50!blue!25!gray}{Countries simultaneously give costly concessions and decide on the efficiency of the concessions: $(g_i, e_i) \in \mathbb{R}_+ \times [0,1]$} 
		\end{enumerate}	
	\item[1--$\infty$.] Countries engage in a simultaneous Prisoners' Dilemma interaction 
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Peace not possible in some scenarios}
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 3}
  When concessions impact the giver's future welfare, peace becomes unachievable for some parameters under which it is achievable when concessions do not impact the giver's future welfare.
\end{beamerboxesrounded}

\pause
\vskip.1in
\begin{itemize}[<+->]
	\item Effect comes in part through change in minimum separating concession
	\item High-type utility may increase or decrease from the benchmark case
\end{itemize}
\end{frame}

\subsection{Add Money Burning}
\begin{frame}{Timeline}
\pause
\begin{enumerate}[<+->]
	\item[-1.] Nature independently determines types of Country 1 and Country 2: $\de_i \in \left\{\de_h,\de_l\right\}$
	\item[0.] Concessions
		\begin{enumerate}
			\item[0a.] Countries decide what proportion of a received concession to invest in civil society (vs. military capabilities): $\alpha_i \in [0,1]$
			\item[0b.] \textcolor{blue}{Countries simultaneously give costly concessions and decide on the efficiency of the concessions: $(g_i, e_i) \in \mathbb{R}_+ \times [0,1]$}
		\end{enumerate}	
	\item[1--$\infty$.] Countries engage in a simultaneous Prisoners' Dilemma interaction 
\end{enumerate}
\end{frame}

\begin{frame}{Burning money is (obviously) unattractive...}
\pause
Now benefit of concession is $eg$

\vskip.1in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Lemma 2}
  If a concessions separating equilibrium exists under no money burning, it is optimal to give efficient gifts ($e=1$) in a concessions separating equilibrium with money burning.
\end{beamerboxesrounded}
\pause
\begin{itemize}[<+->]
	\item The benefit of the gift appears on both sides of the incentive constraint for both individuals, so cancels out
	\item Costs of giving a concession don't change
	\item The benefit appears in the high type's expected utility
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{...but in some cases countries will burn money}
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 4}
  When concessions affect the future welfare of the giver \textit{and} their value can be destroyed, there are parameters under which the optimal equilibrium is a separating equilibrium in which concessions are inefficient.
\end{beamerboxesrounded}
\vskip-.1in
\pause
\[
  U_h=peg - c_h(g) + \textstyle \frac{1}{1-\de_h} \left[pT(1+eg)+(1-p)(W-D(1+eg))\right]
\]
\pause
\vskip-.1in
\begin{itemize}[<+->]
	\item If $p$ is low, concessions likely to be used against you
	\item If $\de_l$ is low, concession has to be large to deter low type from from mimicking high type
\end{itemize}
\end{frame}

\section{Mediation}
\subsection{}
\begin{frame}{Mediator as mechanism designer}
\pause
`Manipulative' mediator: parties report their types, must deliver the stipulated concessions
\pause
\begin{itemize}[<+->]
  \item Mechanism: if two high types, give concession and play `Trust'; Otherwise, no concession and `Fight'
	\item Concession is necessary to get truthful revelation, but only to high type
	\item Need cost of concession for low type to be not too large relative to cost for high type
\end{itemize}
\vskip.1in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 5}
  A mediator restores peace where concessions' future welfare impact destroys it and eliminates inefficient concessions elsewhere.
\end{beamerboxesrounded}

@@

Back to \hyperlink{hurt}{\beamerbutton{Burning money unattractive}}.
\end{frame}
